---
title: "Population of Mumbai"
execute:
  echo: false
  message: false
---

```{r}
#| include: false
library(raster)
library(osmdata)
library(sf)
library(dplyr)
library(ggplot2)
library(patchwork)
library(stars)
library(latex2exp)
```


The Mumbai of 2022 is estimated as a roaring megacity of a population anywhere between 18 and 25 million people @demographiaWorldUrbanAreas2022. A simple population count is difficult to pin down since the [last census was in 2011](https://www.census2011.co.in/census/metropolitan/305-mumbai.html)and due to the pandemic it's [unclear whether the 2021 census has even happened](https://www.census2011.co.in/census/metropolitan/305-mumbai.html) 

Ten years ago, the recorded population of the Mumbai metropolitan region was 18.4 million. The island city recorded a population 12.4 million - with 9.4 million in Mumbai suburban and 3.1 million in Mumbai City. The remaining population is spread across the metropolitan region on mainland - including the nearby exurb planned city turned to urban sprawl called [Navi Mumbai](https://en.wikipedia.org/wiki/Navi_Mumbai) where I grew up. 

My attempts to find a more recent, spatially distributed estimates for the island city (rather than the metropolitan region) led me to the Worldpop estimations (@bondarenkoCensusProjectiondisaggregatedGridded2020a). Some clever statistical models that are calibrated against UNDP counts gives population estimates for Mumbai island in 2020 as 13.38 million people living in Mumbai (@SpatialDistributionPopulation). The total splits 2.8 million to Mumbai city and 10.6 million to Mumbai suburban. These values are a little difficult to reconcile against the census from a decade prior as they note little increase and in the case of the city zone, population appears to have decreased! 

```{r}
#\ include: false

# get population counts from world pop data
pop <- read_stars("~/Documents/worldpop-data/ind_ppp_2020_UNadj_constrained.tif")

# get mumbai polygon for subsetting
mumbai_poly <-  getbb("mumbai india", format_out = "polygon")[[1]] %>% 
  as.data.frame() %>%
  st_as_sf(coords = c("V1", "V2")) %>% 
  # Combine so that whole dataframe becomes one polygon
  st_combine() %>% 
  st_cast("POLYGON") %>% 
  st_set_crs(4326) 

# get mumbai polygon for subsetting
mumbai_suburban_poly <- getbb("mumbai suburban india", format_out = "polygon") %>% 
  as.data.frame() %>%
  st_as_sf(coords = c("V1", "V2")) %>% 
  # Combine so that whole dataframe becomes one polygon
  st_combine() %>% 
  st_cast("POLYGON") %>% 
  st_set_crs(4326) 

mumbai_city_poly <- getbb("mumbai city india", format_out = "polygon")[[1]] %>% 
  as.data.frame() %>%
  st_as_sf(coords = c("V1", "V2")) %>% 
  # Combine so that whole dataframe becomes one polygon
  st_combine() %>% 
  st_cast("POLYGON") %>% 
  st_set_crs(4326) 

```

The worldpop estimates even provide population counts at a 100m$^2$ spatial resolution. Cropping the estimates to the two zones in the island city shows some interesting spatial patterns. For example, Mumbai city appears to be much more homogeneous in population density than Mumbai suburban. 

```{r}
mumbai_pop <- st_crop(pop, mumbai_poly)
mumbai_suburban_pop <- st_crop(pop, mumbai_suburban_poly)
mumbai_city_pop <- st_crop(pop, mumbai_city_poly)

p1 <- ggplot() + 
  geom_stars(data = mumbai_pop) + 
  geom_sf(data = mumbai_poly, size = 1, fill = NA) + 
  scale_fill_viridis_c(limits = c(0, 600), na.value = "transparent") +
  scale_x_continuous(breaks = seq(72.8, 73, by = 0.1)) + 
  theme_minimal() + 
  labs(x = "",
       y = "",
       title = 'Mumbai') + 
  guides(fill = guide_legend("Population")) 

p2 <- ggplot() + 
  geom_stars(data = mumbai_city_pop) + 
  geom_sf(data = mumbai_city_poly, colour = "red", size = 1, fill = NA) + 
  scale_fill_viridis_c(limits = c(0, 600), na.value = "transparent") +
  scale_x_continuous(breaks = seq(72.8, 73, by = 0.1)) + 
  theme_minimal() + 
  labs(x = "",
       y = "",
       title = "City") + 
  theme(legend.position = "None")

p3 <- ggplot() + 
  geom_stars(data = mumbai_suburban_pop) + 
  geom_sf(data = mumbai_suburban_poly, colour = "red", size = 1, fill = NA) + 
  scale_fill_viridis_c(limits = c(0, 600), na.value = "transparent") +
  scale_x_continuous(breaks = seq(72.8, 73, by = 0.1)) + 
  theme_minimal() + 
  labs(x = "",
       y = "", 
       title = "Suburban") + 
  theme(legend.position = "None") 

((p1 | p3 / p2)) + 
  plot_layout(guides = 'collect')

```

```{r}
mumbai_city_pop_df <- mumbai_city_pop %>% 
  as.data.frame() %>% 
  mutate(zone = "Mumbai City")
mumbai_suburban_pop_df <- mumbai_suburban_pop %>% 
  as.data.frame() %>% 
  mutate(zone = "Mumbai Suburban")

mumbai_pop_df <- bind_rows(mumbai_city_pop_df, mumbai_suburban_pop_df)

```

Flattening the spatial population distribution reveals modalities as an overall pattern. In Mumbai City, the older southern part of the island, there are two distinct types of areas: those with population counts of \~250 people /100m$^2$ and those with 50% more density \~375 people / 100m$^2$. The larger zone of Mumbai suburban is more varied containing up to 4 modes - 2 in addition to the ones in Mumbai city. Whether these modes relate to specific urban typologies or if they are an artefact of the population count derivation model is still to be understand. Likely it is both. 

```{r}
#| message: false
#| warning: false
ggplot(mumbai_pop_df) + 
  geom_histogram(aes(x = ind_ppp_2020_UNadj_constrained.tif, 
                     fill = zone),
                 show.legend = FALSE) + 
  facet_wrap(~zone, ncol = 1) + 
  theme_minimal() +
  scale_x_continuous(unname(TeX("Population count per 100$m^2$")),
                   limits = c(0, 600)) + 
  labs(y = "") 
```


